Major Project for Stefan Klaus(\href{mailto:stk4@aber.ac.uk}{\tt stk4@aber.\+ac.\+uk})

Odometry based map building. This project uses a simulator as well as a virtual E-\/\+Puck representation to map different environments.

This project uses the Webots simulator. All code has been created and tested on a Windows system using Webots P\+R\+O 7.\+4.\+3. The project might perform differently on Unix based systems or on older/newer versions of the simulator.

This folder holds all the technical work as well as the final and mid-\/project reports for the project.

An overview over the folders\+:

webots\+: holds the world files for the Webots simulator

Report\+: Holds the source files(\+La\+Te\+X) of the final report as well the project outline specification.

Project data\+: holds sample video files(.mp4) of how the program performs on my system.

Mid Project Presentation\+: holds the power point created for the mid-\/project presentation.

lib\+: holds all source files(besides the main) needed for the project.

figures\+: holds all the figures(including .odg files) and screenshots of the project

doxygen documentation\+: holds the documentation generated for the code, using doxygen.

controllers\+: holds the \hyperlink{_main_8c}{main.\+c} source file for the project. This structure might be confusing however it was used by Webots and that's why I kept it.

Other files\+: just standard files needed for git, the Sublime Text project and workspace file, as well as a copy of the final report for convenience. 